Why robot programming is hard

### 2014-09-04 paper.tex
More refs and rewrote parts of intro section more formally

I never thought I would be a roboticist. In grad school, our robotics team spent months teaching little robot dogs to play soccer and solving other seemingly simple problems. Far better to work in the world of software, I figured, where the systems I built could make a real difference in people's lives. Software could automate repetitive tasks in computer use and free people from mundane work. So instead I developed expertise in human-computer interaction and intelligent user interfaces. My chosen field, end user programming, has the potential to enable millions of ordinary computer users to customize and adapt systems to their own needs.

But the world is changing. Computing has moved off the desktop and into the smartphones, smart environments, and the world around us. Innovation is happening in the physical world, using computing to effect material change in our environment. The past decades have seen industrial robots increase manufacturing efficiency on the factory floor. Even more, we are now seeing robots tackle tasks from our daily lives. Robotic vacuum cleaners are changing the way we clean our homes. Robotic lawnmowers keep the yard tidy without having to lift a finger. Nest's robotic thermostat learns your habits and keeps your home at the optimal temperature. Google's driverless car could change the way we commute.

@@

Yet the holy grail of personal service robots remains elusive. I believe there are three main reasons. First, there is a wide variety of tasks that people want service robots to perform, ranging from folding laundry, to finding lost car keys, to picking up the kids' toys, or doing the dishes and putting them away. Second, these tasks must be performed in a variety of environments, including cluttered homes and kitchens with different affordances such as modern drawer pulls or faucet handles. Finally and most importantly, the robots themselves must be easy to operate by non-technical people.

We have started to see robots enter the marketplace that address two out of those three challenges. For example, iRobot's Roomba\footnote{http://www.irobot.com} only vacuums floors, but it does so in a variety of environments and is simple to use even by non-technical folks. On the other hand, Rethink Robotics' Baxter robot\footnote{http://www.rethinkrobotics.com/products/baxter/} can be easily programmed to perform a variety of tasks in light industrial settings, but its lack of mobility makes it less suitable for the home. As another example, Willow Garage's PR-2 robot has been programmed to perform a variety of household tasks, in a variety of environments, but can only be operated by highly skilled roboticists.

Yet I believe the time is ripe for a breakthrough in personal service robotics, starting with robots that perform simpler tasks in relatively structured environments such as hospitals or hotels, and driven by better design and usable interfaces that enable them to be used by non-expert users.

@@

Yet our experience with Robots for Humanity underscored how difficult it was for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door was very difficult for Henry to accomplish. While Henry had only limited interaction capabilities, he was also strongly motivated because he has no alternatives. Able-bodied consumers, on the other hand, would quickly lose patience and simply fetch the towel themselves. If the state of the art in robot operation is so time-consuming, what hope do we have that ordinary mortals will be able to operate robots to perform these tasks in their own homes?

Granted, designing teleoperation interfaces for motor-impaired users is deliberately trying to solve the hardest problem first.  Yet while Henry's case highlights the challenges of creating usable assistive robotics interfaces, it also illustrates the enormous benefit to humanity of enabling disabled adults to regain some of their independence.

With this goal in mind, personal service robots are more than a consumer toy, but a technology that could changes people's lives for the better.  Why, then, are robots still so difficult to use? What makes robots so difficult to operate?

@@

If you replace computer with robot, we come to a definition of robot programming: finding the right series of instructions that cause a robot to automate a task or exhibit a certain behavior. Every robot makes available a certain set of functionality -- its user-visible API. These are the primitive behaviors that have been programmed into the robot and can be activated by its user. End users must select which behavior to call when, in order to make the robot do what they want.
%Anyone who operates a robot, including consumers who purchase one for their home, will be programming the robot according to this definition.

@@

\item {\bf Task/environment variance}: if the task to be performed is always the same, the programming task is easier than if the goal task changes significantly or if the environment in which the robot must operate is changing. The program must be generalized to handle this variance, which makes the programming process more difficult.
% REF: something from the EUP literature? Andy Ko or Margaret Burnett?

@@

For personal robotics to truly succeed, regular people will need to use robot programming interfaces to tell robots what to do. These interfaces for robot programming will need to be simple enough for ordinary consumers to be able to pick up and use. Until users are able to instruct robots quickly and with low effort, personal robotics will be limited to niche markets. Why is end user robot programming so challenging? Several compounding factors make it more difficult than normal computer use.

@@

%Most robot programs can be simplified by only controlling one moving part at a time: first move the base, then raise the torso, then position the left arm, then position the right arm, finally open the cabinet. For many tasks this is a reasonable solution. However there is a subset of human tasks for which we need multiple parts moving simultaneously, such as picking up a heavy or bulky object with two hands, or turning one's head to look where one is going.

@@

In order to deal with these various types of uncertainty, professional programmers learn to use probabilistic models~\cite{darwiche-cacm10} and craft fault-tolerant code~\cite{wright-cacm09} with frequent error checking. Systems such as Nguyen's ROS Commander~\cite{nguyen-icra13} enable non-roboticists to define and customize robot behaviors, handling error conditions. However, ROS Commander's programming interface, which the authors liken to a Photoshop for robotics, still seems to have a significant learning curve, especially for casual consumers who just want to teach their robot to do something useful in the home.

@@

Reasoning about objects in the world requires some amount of common sense which has yet to be programmed consistently into robots. For example, a robot who has been asked to pick up and deliver a coffee cup may well select a grasp that involves dipping its fingers into the liquid and picking up the cup by its side wall. Although the grasp would be valid, this behavior would be judged somewhat undesireable by human observers.

>>>>

I never thought I would be a roboticist. In grad school I focused on software agents -- {\em softbots}~\cite{etzioni-cacm94} -- that automated repetitive tasks and freed people from mundane work.  I developed expertise in human-computer interaction and intelligent user interfaces. The field of end user programming, where I have focused most of my research, has the potential to enable millions of ordinary computer users to customize and adapt software systems to their own needs.

But the world is changing. Computing has moved off the desktop and into the smartphones, smart environments, and the world around us. Innovation is happening in the physical world, using computing to effect material change in our environment. The past decades have seen industrial robots increase manufacturing efficiency on the factory floor. Even more, we are now seeing robots tackle tasks from our daily lives. Robotic vacuum cleaners are changing the way we clean our homes\footnote{http://www.irobot.com/us/learn/home/roomba}. Robotic lawnmowers keep the yard tidy without having to lift a finger\footnote{http://www.robomow.com}. Nest's robotic thermostat learns your habits and keeps your home at the optimal temperature\footnote{http://www.nest.com}. Google's Self-Driving Car could change the way we commute.

@@

One of the main applications for robotics is to create personal service robots which can assist humans in performing activities of daily living. This vision is particularly attractive as our population ages, and we wish to maintain or increase elders' quality of life despite the rapidly growing costs in the health-care sector~\cite{roy-wire00}.

Yet the goal of creating personal service robots remains elusive. I believe there are three main reasons. First, there is a wide variety of tasks that people want service robots to perform, ranging from folding laundry, to finding lost car keys, to picking up the kids' toys, or doing the dishes and putting them away. Second, these tasks must be performed in a variety of environments, such as cluttered homes and kitchens with different affordances such as modern drawer pulls or faucet handles. Finally and most importantly, the robots themselves must be easy to operate by non-technical people.

We have started to see robots enter the marketplace that address two out of those three challenges. For example, iRobot's Roomba\footnote{http://www.irobot.com} only vacuums floors, but it does so in a variety of environments and is simple to use even by non-technical folks. On the other hand, Rethink Robotics' Baxter robot\footnote{http://www.rethinkrobotics.com/products/baxter/} can be easily programmed to perform a variety of tasks in light industrial settings, but its lack of mobility makes it less suitable for the home. As another example, Willow Garage's PR-2 robot has been programmed to perform a variety of household tasks, in a variety of environments, but can only be operated by highly skilled roboticists~\cite{ciocarlie-iser10}.

I believe the time is ripe for a breakthrough in personal service robotics, starting with robots that perform simpler tasks in relatively structured environments such as hospitals or hotels, and driven by better design and usable interfaces that enable them to be used by non-expert users.

@@

Yet our experience with Robots for Humanity underscored how difficult it was for ordinary mortals to instruct robots what to do. Even using state-of-the-art robot visualization software developed at Willow Garage, simple teleoperation tasks such as opening a cabinet door were very difficult for Henry to accomplish. While Henry had only limited interaction capabilities, he was also strongly motivated because he has no alternatives. Able-bodied consumers, on the other hand, would quickly lose patience and simply fetch the towel themselves. If the state of the art in robot operation is so time-consuming, what hope do we have that ordinary mortals will be able to operate robots to perform these tasks in their own homes?

While Henry's case highlights the challenges of creating usable assistive robotics interfaces, it also illustrates the enormous benefit to humanity of enabling disabled adults to regain some of their independence. With this goal in mind, personal service robots are more than a consumer toy, but a technology that could changes people's lives for the better.  Why, then, are robots still so difficult to use? What makes robots so difficult to operate?

@@

If you replace computer with robot, we come to a definition of robot programming: finding the right series of instructions that cause a robot to automate a task or exhibit a certain behavior. Every robot makes available a certain set of functionality -- its user-visible API. These are the primitive behaviors that have been programmed into the robot and can be activated by its user. End users must select which behavior to call when, in order to make the robot do what they want.

Even the seemingly-simple cleaning functionality of a Roomba vacuum is typically invoked deliberately as part of a larger sequence of housecleaning tasks.  For example, users often reported first doing pre-cleaning tasks such as picking up clutter or setting down obstacles to constrain the robot's path before activating it~\cite{forlizzi-hri06}. Guided by knowledge of the robot's capabilities, these users are selecting when to invoke the robot's functionality in order to have the desired result.  Robots with more functionality only increase the difficulty of this programming process.

@@

\item {\bf Task/environment variance}: if the task to be performed is always the same, the programming task is easier than if the goal task changes significantly or if the environment in which the robot must operate is changing. The program must be generalized to handle this variance, which makes the programming process more difficult~\cite{cypher-chi91,bogart-vlhcc08}.

@@

For personal robotics to truly succeed, regular people will need to use robot programming interfaces to tell robots what to do. These interfaces for robot programming will need to be simple enough for ordinary consumers to be able to pick up and use. Until users are able to instruct robots quickly and with little effort, personal robotics will be limited to niche markets. Why is end user robot programming so challenging? Several compounding factors make it more difficult than normal computer use.

@@

Many robot programs can be simplified by only controlling one moving part at a time: first move the base, then raise the torso, then position the left arm, then position the right arm, finally open the cabinet. For many tasks this is a reasonable solution. However there is a subset of human tasks for which we need multiple parts moving simultaneously, such as picking up a heavy or bulky object with two hands.

@@

In order to deal with these various types of uncertainty, professional programmers learn to use probabilistic models~\cite{darwiche-cacm10} and craft fault-tolerant code~\cite{wright-cacm09} with frequent error checking. Systems such as Nguyen's ROS Commander~\cite{nguyen-icra13} enable non-roboticists to define and customize robot behaviors, including what to do under different error conditions. However, ROS Commander's programming interface, which the authors liken to a Photoshop for robotics, still seems to have a significant learning curve, especially for casual consumers who just want to teach their robot to do something useful in the home.

@@

Reasoning about objects in the world requires some amount of common sense which has yet to be programmed consistently into robots. For example, a robot who has been asked to pick up and deliver a coffee cup will need to be taught to keep the cup upright during transit to avoid spilling its contents.
